# About R {#about-R}

```{r echo = FALSE}
library(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
opts_chunk$set(echo = FALSE) 
knit_hooks$set(inline = function(x) {
        if (!is.numeric(x)) { 
                x 
        } else if (x >= 10000) { 
                prettyNum(round(x, 2), big.mark = ",")
        } else {
                prettyNum(round(x, 2))
        }
})
```

## Features {-}

```{r}
# if internet available, find and save current number of R packages
library("pingr")
if (pingr::is_online()) {
	n_pkg <- nrow(available.packages(contrib.url("https://cloud.r-project.org/")))
	# round down to nearest 250
	n_pkg <- 250 * floor(n_pkg / 250)
	n_pkg <- format(n_pkg, big.mark = ",", trim = TRUE, digits = 0, scientific = FALSE)
	date_pkg <- format(Sys.Date(), "%b %Y")
	info_pkg <- c(n_pkg, date_pkg)
	saveRDS(info_pkg, "data/info_pkg.rds")
}

# use saved values 
info_pkg <- readRDS("data/info_pkg.rds")
n_pkg    <- info_pkg[1]
date_pkg <- info_pkg[2]
```

R is an open source language and environment for statistical computing and graphics [@R-base], ranked by IEEE in 2020 as the 6th most popular programming language (Python, Java, and C are the top three) [@Cass:2020]. If you are new to R, some of its best features, paraphrasing Wickham  [-@wickham2014advanced], are: 

- R is free, open source, and available on every major platform, making it easy for others to replicate your work.
- More than `r n_pkg` open-source R packages are available (`r date_pkg`). Many are cutting-edge tools. 
- R packages provide deep-seated support for data analysis, e.g., missing values, data frames, and subsetting.

RStudio, an integrated development environment (IDE) for R, includes a console, editor, and tools for plotting, history, debugging, and workspace management as well as access to GitHub for collaboration and version control  [@2016rstudio]. 

## Data visualization {-}

Charles Kostelnick [-@kostelnick2007] writes, "The array of design options in software like Microsoft Excel and PowerPoint creates the **illusion** of flexibility. ... So marvelously malleable are these graphical effects---but for whom and to what end? Paradoxically, then, even as the technology for visualizing data has become more sophisticated, it does not necessarily engender rhetorically sensitive design."

The graphics tools in R provide the means to control every pixel in the service of "rhetorically sensitive design." Designers can craft their visual arguments to balance  logos, ethos, pathos, and kairos as appropriate for a given audience in a given rhetorical situation. 

[[For a gallery of R graphics]](https://www.r-graph-gallery.com/){target="_blank"} 

<br>
[&#9650; top of page](#about-R)
