# Working with R {#work-with-R}

```{r echo = FALSE}
library(knitr)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, comment = "#>", collapse = TRUE) 
knitr::knit_hooks$set(inline = function(x) {
        if (!is.numeric(x)) { 
                x 
        } else if (x >= 10000) { 
                prettyNum(round(x, 2), big.mark = ",")
        } else {
                prettyNum(round(x, 2))
        }
})
options(
  datatable.print.nrows = 10,
  datatable.print.topn = 5,
  datatable.print.class = TRUE
)
```

```{r echo = FALSE}
# if internet available, find and save current number of R packages
library("pingr")
if (pingr::is_online()) {
	n_pkg <- nrow(available.packages(contrib.url("https://cloud.r-project.org/")))
	# round down to nearest 250
	n_pkg <- 250 * floor(n_pkg / 250)
	n_pkg <- format(n_pkg, big.mark = ",", trim = TRUE, digits = 0, scientific = FALSE)
	date_pkg <- format(Sys.Date(), "%b %Y")
	info_pkg <- c(n_pkg, date_pkg)
	saveRDS(info_pkg, "data/info_pkg.rds")
}

# use saved values 
info_pkg <- readRDS("data/info_pkg.rds")
n_pkg    <- info_pkg[1]
date_pkg <- info_pkg[2]
```

R is an open source language and environment for statistical computing and graphics [@R-base], ranked by IEEE in 2020 as the 6th most popular programming language (Python, Java, and C are the top three) [@Cass:2020]. If you are new to R, some of its best features, paraphrasing Wickham  [-@wickham2014advanced], are: 

- R is free, open source, and available on every major platform. 
- R packages provide effective tools for data analysis and visualization. 
- More than `r n_pkg` open-source R packages are available (`r date_pkg`). Many are cutting-edge tools. 

RStudio, an integrated development environment (IDE) for R, includes a console, editor, and tools for plotting, history, debugging, and workspace management as well as access to GitHub for collaboration and version control  [@2016rstudio]. 




## Prerequisites {#prerequisites}

Before proceeding, you should have completed

- [Install everything]    
- Launched your workshop project---`workshop.Rproj` or other name that you selected---to start the R session     

With these prerequisites accomplished, and working in your RStudio  `workshop.Rproj` project, you can create a save a script by:

- Use the pulldown menu, _File &gt; New File &gt; R Script_ 
- File menu > Save As... 
- In the dialog box, navigate to your `scripts` directory, type a file name, for example, `01-R-basics.R` (file names in R can start with numerals), and _Save_. 

We suggest you start a new R script for each tutorial and save it to the `scripts` directory. For example, at the end of the workshop, your scripts directory might contain the following files:     

``` r
        \scripts    
            \01-R-basics.R    
            \02-getting-started.R    
            \03-case-study-programs.R    
            \04-case-study-students.R     
            etc.  
```



## New to R? 

[Prerequisites] should be completed before proceeding. By the end of the workshop, our R beginners will have made progress on two or possibly three tutorials:  

- [R basics] An introduction to R. 
- [Getting started](https://midfieldr.github.io/midfieldr/articles/art-000-getting-started.html){target="_blank"}: Examine the MIDFIELD practice data  
- [Case study programs](https://midfieldr.github.io/midfieldr/articles/art-002-case-study-programs.html){target="_blank"} Gather CIP codes and program names

If there is still time remaining, continue to any tutorial listed in the [After the workshop] section.  













## Familiar with R? 

[Prerequisites] should be completed before proceeding. By the end of the workshop, our more experienced R users will have made substantive progress on two or possibly three tutorials: 

- [Getting started](https://midfieldr.github.io/midfieldr/articles/art-000-getting-started.html){target="_blank"}: Examine the MIDFIELD practice data  
- [Case study programs](https://midfieldr.github.io/midfieldr/articles/art-002-case-study-programs.html){target="_blank"} Gather CIP codes and program names 
- [Case study students](https://midfieldr.github.io/midfieldr/articles/art-004-case-study-students.html){target="_blank"} Gather students who pass the data sufficiency criterion. 



If there is still time remaining, continue to any tutorial listed in the [After the workshop] section.  
 
 
 
 
## After the workshop 

At his point, the learning is self-directed. Choose the skills you want to continue working on. We have a set of tutorials for 

- [Developing R skills] 
- [Continuing the case study]
- [Exploring midfieldr functions]





### Developing R skills

The basic skills tutorials take about 50 minutes each. 

- [R basics] 
- [Graph basics]  
- [Data basics]   





### Continuing the case study

The case study is a quick tour of a typical workflow using student unit record data. This is a “big picture” development---functions are used without detailed explanations or development so that we can focus on the logic of the analysis. For exploring midfieldr functions and methods in greater detail, see the vignettes listed in the next section.  

- [Case study programs](https://midfieldr.github.io/midfieldr/articles/art-002-case-study-programs.html){target="_blank"}
- [Case study students](https://midfieldr.github.io/midfieldr/articles/art-004-case-study-students.html){target="_blank"} 
- [Case study stickiness](https://midfieldr.github.io/midfieldr/articles/art-006-case-study-stickiness.html){target="_blank"} 
- [Case study graduation rate](https://midfieldr.github.io/midfieldr/articles/art-008-case-study-grad-rate.html){target="_blank"}





### Exploring midfieldr functions

Deep dive into the midfieldr functionality. The work flow follows the same general pattern as the quicker case study, but pauses to explore each function in more detail, exploring the arguments and strategies for use. In general, each tutorial is self-contained so you may enter at almost any point. 

- [Program codes and names](https://midfieldr.github.io/midfieldr/articles/art-010-program-codes.html){target="_blank"} Practice strategies of searching `cip` for programs we want to study. 
- [Subsetting MIDFIELD data](https://midfieldr.github.io/midfieldr/articles/art-015-subsetting-midfield-data.html){target="_blank"} Use programs codes to subset the MIDFIELD data tables. 
- [Data sufficiency](https://midfieldr.github.io/midfieldr/articles/art-020-data-sufficiency.html){target="_blank"} What it is and how it is applied to student unit-record (SUR) data. 
- [Timely completion](https://midfieldr.github.io/midfieldr/articles/art-030-timely-completion.html){target="_blank"} What it is and how it is applied to SUR data. 
- [FYE programs](https://midfieldr.github.io/midfieldr/articles/art-040-fye-programs.html){target="_blank"} What they are and how they are accommodated with SUR data. 
- [Multiway graphs](https://midfieldr.github.io/midfieldr/articles/art-050-multiway-graphs.html){target="_blank"} How to graph and interpret a common data structure encountered when working with SUR data. 
- [Tabulating data](https://midfieldr.github.io/midfieldr/articles/art-060-tabulating-data.html){target="_blank"} How to tabulate multiway data for publication. 









<br>
[&#9650; top of page](#start-with-R)
