# Working with R {#work-with-R}

```{r echo = FALSE}
library(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
opts_chunk$set(echo = FALSE) 
knit_hooks$set(inline = function(x) {
        if (!is.numeric(x)) { 
                x 
        } else if (x >= 10000) { 
                prettyNum(round(x, 2), big.mark = ",")
        } else {
                prettyNum(round(x, 2))
        }
})
```

```{r}
# if internet available, find and save current number of R packages
library("pingr")
if (pingr::is_online()) {
	n_pkg <- nrow(available.packages(contrib.url("https://cloud.r-project.org/")))
	# round down to nearest 250
	n_pkg <- 250 * floor(n_pkg / 250)
	n_pkg <- format(n_pkg, big.mark = ",", trim = TRUE, digits = 0, scientific = FALSE)
	date_pkg <- format(Sys.Date(), "%b %Y")
	info_pkg <- c(n_pkg, date_pkg)
	saveRDS(info_pkg, "data/info_pkg.rds")
}

# use saved values 
info_pkg <- readRDS("data/info_pkg.rds")
n_pkg    <- info_pkg[1]
date_pkg <- info_pkg[2]
```

R is an open source language and environment for statistical computing and graphics [@R-base], ranked by IEEE in 2020 as the 6th most popular programming language (Python, Java, and C are the top three) [@Cass:2020]. If you are new to R, some of its best features, paraphrasing Wickham  [-@wickham2014advanced], are: 

- R is free, open source, and available on every major platform, making it easy for others to replicate your work.
- More than `r n_pkg` open-source R packages are available (`r date_pkg`). Many are cutting-edge tools. 
- R packages provide deep-seated support for data analysis, e.g., missing values, data frames, and subsetting.

RStudio, an integrated development environment (IDE) for R, includes a console, editor, and tools for plotting, history, debugging, and workspace management as well as access to GitHub for collaboration and version control  [@2016rstudio]. 

## New to R? 

- [R basics](https://midfieldr.github.io/workshops/r-basics.html) from the 2019 MIDFIELD Institute (editing required)
- [Introduction to midfieldr](https://midfieldr.github.io/midfieldr/articles/art-00-midfield-data.html): a midfieldr vignette 
- [Multiway graphs and tables](https://midfieldr.github.io/midfieldr/articles/art-01-multiway.html): a midfieldr vignette 


## New to midfieldr? 

- [Introduction to midfieldr](https://midfieldr.github.io/midfieldr/articles/art-00-midfield-data.html): a midfieldr vignette 
- [Multiway graphs and tables](https://midfieldr.github.io/midfieldr/articles/art-01-multiway.html): a midfieldr vignette 
- [Program codes](https://midfieldr.github.io/midfieldr/articles/art-02-program-codes.html): a midfieldr vignette  

## Additional material for beginners 

- [Graph basics](https://midfieldr.github.io/workshops/graph-basics.html) from the 2019 MIDFIELD Institute (editing required)
- [Data basics](https://midfieldr.github.io/workshops/data-basics.html) from the 2019 MIDFIELD Institute (editing required) 

## Additional midfieldr tutorials

- [Case selection](https://midfieldr.github.io/midfieldr/articles/art-03-case-selection.html): a midfieldr vignette  
- [Data sufficiency](https://midfieldr.github.io/midfieldr/articles/art-04-data-sufficiency.html): a midfieldr vignette  
- [Timely completion](https://midfieldr.github.io/midfieldr/articles/art-06-timely-completion.html): a midfieldr vignette  
- [Stickiness](https://midfieldr.github.io/midfieldr/articles/art-05-stickiness.html): a midfieldr vignette 
- [FYE programs](https://midfieldr.github.io/midfieldr/articles/art-07-fye-programs.html): a midfieldr vignette  
- [Graduation rate](https://midfieldr.github.io/midfieldr/articles/art-08-graduation-rate.html): a midfieldr vignette  







<br>
[&#9650; top of page](#start-with-R)
