# Data basics {#data-basics}

```{r echo = FALSE}
library(knitr)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, comment = "#>", collapse = TRUE) 
knitr::knit_hooks$set(inline = function(x) {
        if (!is.numeric(x)) { 
                x 
        } else if (x >= 10000) { 
                prettyNum(round(x, 2), big.mark = ",")
        } else {
                prettyNum(round(x, 2))
        }
})
options(
  datatable.print.nrows = 10,
  datatable.print.topn = 5,
  datatable.print.class = TRUE
)
```

## Introduction 

This tutorial is an introduction to ...


If you already have R experience, you might still want to browse this section in case you find something new. 

[Prerequisites] should be completed before proceeding. After that, the tutorial should take about an hour. 

Create a new script for this tutorial. 

- See [Open an R script] if you need a refresher on creating, saving, and running an R script. 
- At the top of the script add a minimal header and install and load the packages indicated.  

```{r}
# Data basics 
# name 
# date 

library("midfieldr")
library("data.table")
library("ggplot2")
library("gapminder")

# Optional code to control data.table printing
options(
  datatable.print.nrows = 10,
  datatable.print.topn = 5,
  datatable.print.class = TRUE
)
```

If you get an error like this one after running the script, 

        Error in library("gapminder") : there is no package called 'gapminder'

then the package needs to be installed. If you need a refresher on installing packages, see [Install CRAN packages]. Once the missing package is installed, you can rerun the script. 

Guidelines 

- As you work through the tutorial, type a line or chunk of code then *File &gt; Save* and run the script.    
- Confirm that your result matches the tutorial result.  
- **Your turn** exercises give you chance to devise your own examples and check them out. You learn by doing (but you knew that already)!
 





<br>
[&#9650; top of page](#start-with-R)
