# Working with R {#work-with-R}

```{r echo = FALSE}
library(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
opts_chunk$set(echo = FALSE) 
knit_hooks$set(inline = function(x) {
        if (!is.numeric(x)) { 
                x 
        } else if (x >= 10000) { 
                prettyNum(round(x, 2), big.mark = ",")
        } else {
                prettyNum(round(x, 2))
        }
})
```

```{r}
# if internet available, find and save current number of R packages
library("pingr")
if (pingr::is_online()) {
	n_pkg <- nrow(available.packages(contrib.url("https://cloud.r-project.org/")))
	# round down to nearest 250
	n_pkg <- 250 * floor(n_pkg / 250)
	n_pkg <- format(n_pkg, big.mark = ",", trim = TRUE, digits = 0, scientific = FALSE)
	date_pkg <- format(Sys.Date(), "%b %Y")
	info_pkg <- c(n_pkg, date_pkg)
	saveRDS(info_pkg, "data/info_pkg.rds")
}

# use saved values 
info_pkg <- readRDS("data/info_pkg.rds")
n_pkg    <- info_pkg[1]
date_pkg <- info_pkg[2]
```

R is an open source language and environment for statistical computing and graphics [@R-base], ranked by IEEE in 2020 as the 6th most popular programming language (Python, Java, and C are the top three) [@Cass:2020]. If you are new to R, some of its best features, paraphrasing Wickham  [-@wickham2014advanced], are: 

- R is free, open source, and available on every major platform. 
- R packages provide effective tools for data analysis and visualization. 
- More than `r n_pkg` open-source R packages are available (`r date_pkg`). Many are cutting-edge tools. 

RStudio, an integrated development environment (IDE) for R, includes a console, editor, and tools for plotting, history, debugging, and workspace management as well as access to GitHub for collaboration and version control  [@2016rstudio]. 






## New to R? 

This is the workshop tutorial entry point for the R beginner. Before proceeding, you should have completed

- [Install everything]
- [Set up a project]


Launched your workshop project (`workshop.Rproj` or the name you used) to start the R session 

By the end of the workshop, our R beginners will have made substantive progress on three tutorials: 

- [R basics] A short introduction to R (in the next section) 
- [Getting started](https://midfieldr.github.io/midfieldr/articles/art-00-midfield-data.html): An introduction to the MIDFIELD practice data tables  
- [Multiway graphs and tables](https://midfieldr.github.io/midfieldr/articles/art-01-multiway.html):  Practice working with a data structure commonly encountered when preparing graphs and tables.  












## Familiar with R? 

This is the entry point for participants who are familiar with R. Reminder: You should have completed

- [Install everything]
- [Set up a project]

By the end of the workshop, our more experienced R users will have made substantive progress on three tutorials: 

- [Getting started](https://midfieldr.github.io/midfieldr/articles/art-00-midfield-data.html): An introduction to the MIDFIELD practice data tables  
- [Multiway graphs and tables](https://midfieldr.github.io/midfieldr/articles/art-01-multiway.html):  Practice working with a data structure commonly encountered when preparing graphs and tables.
- [Program codes](https://midfieldr.github.io/midfieldr/articles/art-02-program-codes.html): Explore strategies of searching CIP codes for programs you want to study. 

## After the workshop 

For additional basic skills in R: 

- [Graph basics] 
- [Data basics]  

To continue the midfieldr vignette sequence:  

- [Case selection](https://midfieldr.github.io/midfieldr/articles/art-03-case-selection.html): Select and name a specific set of programs to explore. 
- [Data sufficiency](https://midfieldr.github.io/midfieldr/articles/art-04-data-sufficiency.html): Limited a study to students whose timely completion term is within the range of data provided by their institution.  
- [Timely completion](https://midfieldr.github.io/midfieldr/articles/art-06-timely-completion.html): Graduates completing their program later than their timely completion term are grouped with non-graduates for purposes of computing a persistence metric.  
- [Stickiness](https://midfieldr.github.io/midfieldr/articles/art-05-stickiness.html): Compute longitudinal “stickiness”, the ratio of the number of students graduating in a program to the number of students ever enrolled in that program.  
- [FYE programs](https://midfieldr.github.io/midfieldr/articles/art-07-fye-programs.html): Predict the engineering program the FYE student would have declared had they not been required to enroll in FYE.  
- [Graduation rate](https://midfieldr.github.io/midfieldr/articles/art-08-graduation-rate.html): Compute graduation rate with a more inclusive definition.  



## R basics

text


### Style guide {#style-guide}

### Everything in R has a name 

### Everything in R is an object 

### Use functions to do things

### R functions come in packages

### R objects have class

### R objects have structure

### R only does what you tell it

### Keyboard shortcuts

### What's next?



## Graph basics


## Data basics









<br>
[&#9650; top of page](#start-with-R)
